Exact Phylodynamics via Structured Markov Genealogy Processes

### 2024-02-22 ms.Rnw
corrections

Application of these rules at each jump of $\Xr_t$ constructs a chain of genealogies $\Grh_k$.
In particular, at each jump-time $\Trh_k$, the genealogy $\Grh_{k-1}$ is modified according to the jump-mark $\Urh_k$ to yield $\Grh_k$.
We view $\Grh_k$ as the embedded chain of the continuous-time genealogy process $\Gr_t$.
It is very important to note that, conditional on $\Urh_k$, the number of parents and number of offspring in each deme is determined and the random choice of which lineages die, migrate, are sampled, or sire offspring is independent of the choices at any other times and independent of $(\Trh_k,\Xrh_k)$ for all $k$.
Moreover, by construction, any lineage within a deme is as likely as any other lineage in that deme to be selected as a parent or for death, sampling, or migration.
We refer to this property as the \emph{exchangeability} of lineages within a deme.

@@

@@

@@

\begin{lemma}
  \label{lemma:sing-filt}
  Suppose $\mathcal{M}\subset{\Rp}$ is finite and let $\mathcal{M}_t\coloneq\mathcal{M}\cap{[0,t]}$.
  Suppose that $B:\Rp\times\Xspace^2\to\Rp$ is measurable and
  $\Vr_t$ is an $\Rp$-valued process satisfying
  \begin{equation*}
    \CondExpect{\Vr_t}{\Hr_t=\H_t}=\Indicator{\event{\H}_{t}\supseteq\mathcal{M}_t}\,\prod_{e\in\event{\H}_{t}}{B(e,\Xt_e,\X_e)}.
  \end{equation*}
  Then $\CondExpect{\Vr_t}{\Xr_t=x}$ is a measure-valued process on $\Rp^{\vert\mathcal{M}_t\vert}\times\Xspace$ with density $w(t,x)$ that satisfies
  \begin{equation}
    \label{eq:two-part-filter}
    \begin{aligned}
      \frac{\partial{w}}{\partial{t}}(t,x)&=
      \int{w(t,x')\,\alpha(t,x',x)\,B(t,x',x)\,\dd{x'}}-\int{w(t,x)\,\alpha(t,x,x')\,\dd{x'}},\qquad
      &t\notin{\mathcal{M}},\\
      w(t,x)&=\int{\wt(t,x')\,\alpha(t,x',x)\,B(t,x',x)\,\dd{x'}},\qquad
      &t\in{\mathcal{M}}.
    \end{aligned}
  \end{equation}
  Moreover, $\Expect{\Vr_t}=\int{w(t,x)\,\dd{x}}$.
\end{lemma}

>>>>

Application of these rules at each jump of $\Xr_t$ constructs a chain of genealogies $\Grh_k$.
In particular, at each jump-time $\Trh_k$, the genealogy $\Grh_{k-1}$ is modified according to the jump-mark $\Urh_k$ to yield $\Grh_k$.
We view $\Grh_k$ as the embedded chain of the continuous-time genealogy process $\Gr_t$.
It is very important to note that, conditional on $(\Xrh_k,\Urh_k)$, the number of parents and number of offspring in each deme is determined and the random choice of which lineages die, migrate, are sampled, or sire offspring is independent of these choices at any other times and independent of $(\Xrh_j,\Urh_j)$ for all $j\ne{k}$.
Moreover, by construction, any lineage within a deme is as likely as any other lineage in that deme to be selected as a parent or for death, sampling, or migration.
We refer to this property as the \emph{exchangeability} of lineages within a deme.
Finally, note that $\Gr_t$ does not have the Markov property, though $(\Xr_t,\Ur_t,\Gr_t)$ and $(\Xr_t,\Gr_t)$ do.

@@

\begin{remark}
  Note that, since $\func{Y}_t(\Z)$ is finite, it is permissible to choose $q$ and $\pi$ to be uniform.
\end{remark}

@@

\AAK{[We should define the singular filter equations (\ie those with a discrete portion) also.]}

@@

\begin{lemma}
  \label{lemma:sing-filt}
  Suppose $\mathcal{M}\subset{\Rp}$ is countable such that $\mathcal{M}_t\coloneq\mathcal{M}\cap{[0,t]}$ is finite for all $t$.
  Suppose that $B:\Rp\times\Xspace^2\to\Rp$ is measurable and
  $\Vr_t$ is an $\Rp$-valued process satisfying
  \begin{equation*}
    \CondExpect{\Vr_t}{\Hr_t=\H_t}=\Indicator{\event{\H}_{t}\supseteq\mathcal{M}_t}\,\prod_{e\in\event{\H}_{t}}{B(e,\Xt_e,\X_e)}.
  \end{equation*}
  Then $\CondExpect{\Vr_t}{\Xr_t=x}$ is a measure-valued process on $\Rp^{\vert\mathcal{M}_t\vert}\times\Xspace$ with density $w(t,x)$ that satisfies
  \begin{equation}
    \label{eq:two-part-filter}
    \begin{aligned}
      \frac{\partial{w}}{\partial{t}}(t,x)&=
      \int{w(t,x')\,\alpha(t,x',x)\,B(t,x',x)\,\dd{x'}}-\int{w(t,x)\,\alpha(t,x,x')\,\dd{x'}},\qquad
      &t\notin{\mathcal{M}},\\
      w(t,x)&=\int{\wt(t,x')\,\alpha(t,x',x)\,B(t,x',x)\,\dd{x'}},\qquad
      &t\in{\mathcal{M}}.
    \end{aligned}
  \end{equation}
  Moreover, $\Expect{\Vr_t}=\int{w(t,x)\,\dd{x}}$.
\end{lemma}
